---
title: "RMNP Intelimon Update"
author: "Christina Fossum"
date: "`r Sys.Date()`"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = "C:/Users/cfossum/OneDrive - DOI/Desktop/TLS repository")
```

# Models for RMNP intelimon data

## Contents:
1. ROMO_East_Fuels:
    a. ROMO_East_Fuels_PRE_TotalAll: model for pre-burn total fuel load (tons/acre)
    b. ROMO_East_Fuels_POST_TotalAll: model for post-burn total fuel load (tons/acre)
    c. ROMO_East_Fuels_YR01_TotalAll: model for yr01-post-burn total fuel load (tons/acre)
    
2. ROMO_East_Trees
    a. ROMO_East_TotalTreeCount: model for total tree count per plot
    b. ROMO_East_LiveTreeCount: model for live tree count per plot
    c. ROMO_East_meanCBH: model for mean canopy base height

3. ROMO_East_RX
    a. ROMO_East_RX_SubSev: model for post-burn substate burn severity
    b. ROMO_East_RX_MeanScPct: model for average percent scorch
    c. ROMO_East_RX_MeanScHt: model for average scorch height (m)

4. ROMO_West_Fuels
    a. ROMO_West_Fuels_TotalAll: model for mixed conifer total fuel load (tons/acre)
    b. ROMO_West_Fuels_TotalWood: model for mixed conifer total woody fuel load (tons/acre)
    





# 1. ROMO_East_Fuels

## a. ROMO_East_Fuels_PRE_TotalAll

model formula:
lm1 <- lm(formula = train_pred$TotalAll ~  hr100_1000_l1_median + hr100_1000_l1_tgi)

other top model variables:
- hr0_10_l1_vari
- s_lng_prop_sk
- maxSD
- h_l2_median
- h_US_cnt

```{r}
library(tidyverse)
library(leaps)
library(performance)
library(ggplot2)

# Load pred / vars data
pred <- read.csv("TLS/intelimon models/March 2026 Intelimon Update/RMNP Intelimon Update/1. ROMO_East_Fuels/a. ROMO_East_Fuels_PRE_TotalAll/ROMO_East_Fuels_PRE_TotalAll_pred.csv")
vars <- read.csv("TLS/intelimon models/March 2026 Intelimon Update/RMNP Intelimon Update/1. ROMO_East_Fuels/a. ROMO_East_Fuels_PRE_TotalAll/ROMO_East_Fuels_PRE_TotalAll_vars.csv")

# Create training/test data
set.seed(123)
sample_size <- floor(0.2* nrow(vars))
test_indices <- sample(seq_len(nrow(vars)), size = sample_size)
train_vars <- vars[-test_indices, ]
test_vars <- vars[test_indices, ]
train_pred <- pred[-test_indices, ]
test_pred <- pred[test_indices, ]

# Load model
lm1<- readRDS("TLS/intelimon models/March 2026 Intelimon Update/RMNP Intelimon Update/1. ROMO_East_Fuels/a. ROMO_East_Fuels_PRE_TotalAll/ROMO_East_Fuels_PRE_TotalAll.rda")

#look at the summary r2 values and RMSE & check model for colinearity
summary(lm1)
r2(lm1) 
rmse(lm1)
check_model(lm1) 

# Plot observed values vs. predicted
lm_out <- predict.lm(lm1, vars)
data<-data.frame(x=lm_out, y=pred$TotalAll)

ggplot(data,aes(x,y)) +
  geom_point() +
  geom_smooth(method='lm', se=FALSE,) +
  theme_light() +
  labs(x='Predicted Values', y='Observed Values', title='Pre-Burn Total Fuels') +
  theme(plot.title = element_text(hjust=0.5, size=20, face='bold') 
  ) 


# use the model to predict test data
test_out<- predict.lm(lm1,test_vars)
errors<- test_out - test_pred$TotalAll
mae <- mean(abs(errors))
mse <- mean(errors^2)
rmse <- sqrt(mse)
rmse 

# Plot test observed values vs. test predicted
data2<- data.frame(x=test_out, y=test_pred$TotalAll)
ggplot(data2,aes(x,y)) +
  geom_point() +
  geom_smooth(method='lm', se=FALSE,) +
  theme_light() +
  labs(x='Predicted Values', y='Observed Values', title='test') +
  theme(plot.title = element_text(hjust=0.5, size=20, face='bold') 
  ) 



```


## b. ROMO_East_Fuels_POST_TotalAll

model formula:
lm1 <- lm(formula = train_pred$TotalAll ~ h_l1_kurt    +  h_zpcum4    +   LF_SLPD )

other top model variables:
- Basalarea, maxTH, CBH
- LF_FBFM13
- h_zpcum1, h_zq5, h_zpcum5
- OLAI
- fuel0_3l1_vari, fuel0_3l1_median


```{r}
library(tidyverse)
library(leaps)
library(performance)
library(ggplot2)

# Load pred / vars data

pred <- read.csv("TLS/intelimon models/March 2026 Intelimon Update/RMNP Intelimon Update/1. ROMO_East_Fuels/b. ROMO_East_Fuels_POST_TotalAll/ROMO_East_Fuels_POST_TotalAll_pred.csv")
vars <- read.csv("TLS/intelimon models/March 2026 Intelimon Update/RMNP Intelimon Update/1. ROMO_East_Fuels/b. ROMO_East_Fuels_POST_TotalAll/ROMO_East_Fuels_POST_TotalAll_vars.csv")

# Create training/test data
set.seed(123)
sample_size <- floor(0.2* nrow(vars))
test_indices <- sample(seq_len(nrow(vars)), size = sample_size)
train_vars <- vars[-test_indices, ]
test_vars <- vars[test_indices, ]
train_pred <- pred[-test_indices, ]
test_pred <- pred[test_indices, ]


# Load model
lm1<- readRDS("TLS/intelimon models/March 2026 Intelimon Update/RMNP Intelimon Update/1. ROMO_East_Fuels/b. ROMO_East_Fuels_POST_TotalAll/ROMO_East_Fuels_POST_TotalAll.rda")

#look at the summary r2 values and RMSE & check model for colinearity
summary(lm1)
r2(lm1) 
rmse(lm1)
check_model(lm1) 

# Plot observed values vs. predicted
lm_out <- predict.lm(lm1, vars)
data<-data.frame(x=lm_out, y=pred$TotalAll)

ggplot(data,aes(x,y)) +
  geom_point() +
  geom_smooth(method='lm', se=FALSE,) +
  theme_light() +
  labs(x='Predicted Values', y='Observed Values', title='Post-Burn Total Fuels') +
  theme(plot.title = element_text(hjust=0.5, size=20, face='bold') 
  ) 


# use the model to predict test data
test_out<- predict.lm(lm1,test_vars)
errors<- test_out - test_pred$TotalAll
mae <- mean(abs(errors))
mse <- mean(errors^2)
rmse <- sqrt(mse)
rmse 

# Plot test observed values vs. test predicted
data2<- data.frame(x=test_out, y=test_pred$TotalAll)
ggplot(data2,aes(x,y)) +
  geom_point() +
  geom_smooth(method='lm', se=FALSE,) +
  theme_light() +
  labs(x='Predicted Values', y='Observed Values', title='test') +
  theme(plot.title = element_text(hjust=0.5, size=20, face='bold') 
  ) 



```



## c. ROMO_East_Fuels_YR01_TotalAll

model formula:
lm1 <- lm(formula = train_pred$TotalAll ~  h_ground_cnt + shrubs_l1_median )

other top model variables:
- s_lng_zero_per
- shrubs_l1_skew
- LAI
- hr100_1000_l1_vari

```{r}
library(tidyverse)
library(leaps)
library(performance)
library(ggplot2)

# Load pred / vars data
pred <- read.csv("TLS/intelimon models/March 2026 Intelimon Update/RMNP Intelimon Update/1. ROMO_East_Fuels/c. ROMO_East_Fuels_YR01_TotalAll/ROMO_East_Fuels_YR01_TotalAll_pred.csv")
vars <- read.csv("TLS/intelimon models/March 2026 Intelimon Update/RMNP Intelimon Update/1. ROMO_East_Fuels/c. ROMO_East_Fuels_YR01_TotalAll/ROMO_East_Fuels_YR01_TotalAll_vars.csv")

# Create training/test data
set.seed(123)
sample_size <- floor(0.25* nrow(vars))
test_indices <- sample(seq_len(nrow(vars)), size = sample_size)
train_vars <- vars[-test_indices, ]
test_vars <- vars[test_indices, ]
train_pred <- pred[-test_indices, ]
test_pred <- pred[test_indices, ]

# Load model
lm1<- readRDS("TLS/intelimon models/March 2026 Intelimon Update/RMNP Intelimon Update/1. ROMO_East_Fuels/c. ROMO_East_Fuels_YR01_TotalAll/ROMO_East_Fuels_YR01_TotalAll.rda")

#look at the summary r2 values and RMSE & check model for colinearity
summary(lm1)
r2(lm1) 
rmse(lm1)
check_model(lm1) 

# Plot observed values vs. predicted
lm_out <- predict.lm(lm1, vars)
data<-data.frame(x=lm_out, y=pred$TotalAll)

ggplot(data,aes(x,y)) +
  geom_point() +
  geom_smooth(method='lm', se=FALSE,) +
  theme_light() +
  labs(x='Predicted Values', y='Observed Values', title='Yr01 Post-Burn Total Fuels') +
  theme(plot.title = element_text(hjust=0.5, size=20, face='bold') 
  ) 


# use the model to predict test data
test_out<- predict.lm(lm1,test_vars)
errors<- test_out - test_pred$TotalAll
mae <- mean(abs(errors))
mse <- mean(errors^2)
rmse <- sqrt(mse)
rmse 

# Plot test observed values vs. test predicted
data2<- data.frame(x=test_out, y=test_pred$TotalAll)
ggplot(data2,aes(x,y)) +
  geom_point() +
  geom_smooth(method='lm', se=FALSE,) +
  theme_light() +
  labs(x='Predicted Values', y='Observed Values', title='test') +
  theme(plot.title = element_text(hjust=0.5, size=20, face='bold') 
  ) 



```






# 2. ROMO_East_Trees

## a. ROMO_East_TotalTreeCount

model formula:
lm1 <- lm(formula = train_pred$TreeCount ~ h_MS_cnt + h_MS_per + fuel0_3l1_kurt)

other top model variables:
- h_zq20

```{r}
library(tidyverse)
library(leaps)
library(performance)
library(ggplot2)

# Load pred / vars data
pred <- read.csv("TLS/intelimon models/March 2026 Intelimon Update/RMNP Intelimon Update/2. ROMO_East_Trees/a. ROMO_East_TotalTreeCount/ROMO_EAST_TotalTreeCount_pred.csv")
vars <- read.csv("TLS/intelimon models/March 2026 Intelimon Update/RMNP Intelimon Update/2. ROMO_East_Trees/a. ROMO_East_TotalTreeCount/ROMO_EAST_TotalTreeCount_vars.csv")

# Create training/test data
set.seed(123)
sample_size <- floor(0.2* nrow(vars))
test_indices <- sample(seq_len(nrow(vars)), size = sample_size)
train_vars <- vars[-test_indices, ]
test_vars <- vars[test_indices, ]
train_pred <- pred[-test_indices, ]
test_pred <- pred[test_indices, ]

# Load Model
lm1 <- readRDS("TLS/intelimon models/March 2026 Intelimon Update/RMNP Intelimon Update/2. ROMO_East_Trees/a. ROMO_East_TotalTreeCount/ROMO_EAST_TotalTreeCount.rda")


#look at the summary r2 values and RMSE & check model for colinearity
summary(lm1)
r2(lm1) 
rmse(lm1)
check_model(lm1) 

# Plot observed values vs. predicted
lm_out <- predict.lm(lm1, vars)
data<-data.frame(x=lm_out, y=pred$TreeCount)

ggplot(data,aes(x,y)) +
  geom_point() +
  geom_smooth(method='lm', se=FALSE,) +
  theme_light() +
  labs(x='Predicted Values', y='Observed Values', title='Total Tree Count') +
  theme(plot.title = element_text(hjust=0.5, size=20, face='bold') 
  ) 


# use the model to predict test data
test_out<- predict.lm(lm1,test_vars)
errors<- test_out - test_pred$TreeCount
mae <- mean(abs(errors))
mse <- mean(errors^2)
rmse <- sqrt(mse)
rmse 

# Plot test observed values vs. test predicted
data2<- data.frame(x=test_out, y=test_pred$TreeCount)
ggplot(data2,aes(x,y)) +
  geom_point() +
  geom_smooth(method='lm', se=FALSE,) +
  theme_light() +
  labs(x='Predicted Values', y='Observed Values', title='test') +
  theme(plot.title = element_text(hjust=0.5, size=20, face='bold') 
  ) 


```


## b. ROMO_East_LiveTreeCount

model formula: 
lm1 <- lm(formula = train_pred$L ~ h_l2_per  +    h_l5_cnt + fine_l1_vari)

Other top model variables:
- CanopyCover
- fuel0_3l1_cnt

```{r}
library(tidyverse)
library(leaps)
library(performance)
library(ggplot2)

# Load pred / vars data
pred <- read.csv("TLS/intelimon models/March 2026 Intelimon Update/RMNP Intelimon Update/2. ROMO_East_Trees/b. ROMO_East_LiveTreeCount/ROMO_EAST_LiveTreeCount_pred.csv")
vars <- read.csv("TLS/intelimon models/March 2026 Intelimon Update/RMNP Intelimon Update/2. ROMO_East_Trees/b. ROMO_East_LiveTreeCount/ROMO_EAST_LiveTreeCount_vars.csv")

# Create training/test data
set.seed(123)
sample_size <- floor(0.2* nrow(vars))
test_indices <- sample(seq_len(nrow(vars)), size = sample_size)
train_vars <- vars[-test_indices, ]
test_vars <- vars[test_indices, ]
train_pred <- pred[-test_indices, ]
test_pred <- pred[test_indices, ]

# Load Model
lm1 <- readRDS("TLS/intelimon models/March 2026 Intelimon Update/RMNP Intelimon Update/2. ROMO_East_Trees/b. ROMO_East_LiveTreeCount/ROMO_EAST_LiveTreeCount.rda")


#look at the summary r2 values and RMSE & check model for colinearity
summary(lm1)
r2(lm1) 
rmse(lm1)
check_model(lm1) 

# Plot observed values vs. predicted
lm_out <- predict.lm(lm1, vars)
data<-data.frame(x=lm_out, y=pred$L)

ggplot(data,aes(x,y)) +
  geom_point() +
  geom_smooth(method='lm', se=FALSE,) +
  theme_light() +
  labs(x='Predicted Values', y='Observed Values', title='Live Tree Count') +
  theme(plot.title = element_text(hjust=0.5, size=20, face='bold') 
  ) 


# use the model to predict test data
test_out<- predict.lm(lm1,test_vars)
errors<- test_out - test_pred$L
mae <- mean(abs(errors))
mse <- mean(errors^2)
rmse <- sqrt(mse)
rmse 

# Plot test observed values vs. test predicted
data2<- data.frame(x=test_out, y=test_pred$L)
ggplot(data2,aes(x,y)) +
  geom_point() +
  geom_smooth(method='lm', se=FALSE,) +
  theme_light() +
  labs(x='Predicted Values', y='Observed Values', title='test') +
  theme(plot.title = element_text(hjust=0.5, size=20, face='bold') 
  ) 


```


## c. ROMO_East_meanCBH

model formula:
lm1 <- lm(formula = train_pred$meanCBH ~   LAHV + MeanSA + hr0_10_l1_std)

Other top model variables:
- SDSHT
- MLAI
- h_ng_vari
- vox_l1_vari
- shrubs_l1_tgi

```{r}
library(tidyverse)
library(leaps)
library(performance)
library(ggplot2)

# Load pred / vars data
pred <- read.csv("TLS/intelimon models/March 2026 Intelimon Update/RMNP Intelimon Update/2. ROMO_East_Trees/c. ROMO_East_meanCBH/ROMO_EAST_meanCBH_pred.csv")
vars <- read.csv("TLS/intelimon models/March 2026 Intelimon Update/RMNP Intelimon Update/2. ROMO_East_Trees/c. ROMO_East_meanCBH/ROMO_EAST_meanCBH_vars.csv")

# Create training/test data
set.seed(123)
sample_size <- floor(0.1* nrow(vars))
test_indices <- sample(seq_len(nrow(vars)), size = sample_size)
train_vars <- vars[-test_indices, ]
test_vars <- vars[test_indices, ]
train_pred <- pred[-test_indices, ]
test_pred <- pred[test_indices, ]

# Load Model
lm1 <- readRDS("TLS/intelimon models/March 2026 Intelimon Update/RMNP Intelimon Update/2. ROMO_East_Trees/c. ROMO_East_meanCBH/ROMO_EAST_meanCBH.rda")


#look at the summary r2 values and RMSE & check model for colinearity
summary(lm1)
r2(lm1) 
rmse(lm1)
check_model(lm1) 

# Plot observed values vs. predicted
lm_out <- predict.lm(lm1, vars)
data<-data.frame(x=lm_out, y=pred$meanCBH)

ggplot(data,aes(x,y)) +
  geom_point() +
  geom_smooth(method='lm', se=FALSE,) +
  theme_light() +
  labs(x='Predicted Values', y='Observed Values', title='Canopy Base Height') +
  theme(plot.title = element_text(hjust=0.5, size=20, face='bold') 
  ) 


# use the model to predict test data
test_out<- predict.lm(lm1,test_vars)
errors<- test_out - test_pred$meanCBH
mae <- mean(abs(errors))
mse <- mean(errors^2)
rmse <- sqrt(mse)
rmse 

# Plot test observed values vs. test predicted
data2<- data.frame(x=test_out, y=test_pred$meanCBH)
ggplot(data2,aes(x,y)) +
  geom_point() +
  geom_smooth(method='lm', se=FALSE,) +
  theme_light() +
  labs(x='Predicted Values', y='Observed Values', title='test') +
  theme(plot.title = element_text(hjust=0.5, size=20, face='bold') 
  ) 


```






# 3. ROMO_East_RX

## a. ROMO_East_RX_SubSev

model formula:
lm1 <- lm(formula = train_pred$SubSev ~    hr0_10_l1_median + s_l2_zero_per + h_MS_kurt )

Other top model variables:
- s_l4_prop_ku
- h_l4_skew, h_l4_per
- h_US_tgi, h_US_median
- h_l2_std
- s_l1_prop_ku

```{r}
library(tidyverse)
library(leaps)
library(performance)
library(ggplot2)

# Load pred / vars data
pred <- read.csv("TLS/intelimon models/March 2026 Intelimon Update/RMNP Intelimon Update/3. ROMO_East_RX/a. ROMO_East_RX_SubSev/ROMO_East_RX_SubSev_pred.csv")
vars <- read.csv("TLS/intelimon models/March 2026 Intelimon Update/RMNP Intelimon Update/3. ROMO_East_RX/a. ROMO_East_RX_SubSev/ROMO_East_RX_SubSev_vars.csv")

# Create training/test data
set.seed(123)
sample_size <- floor(0.3* nrow(vars))
test_indices <- sample(seq_len(nrow(vars)), size = sample_size)
train_vars <- vars[-test_indices, ]
test_vars <- vars[test_indices, ]
train_pred <- pred[-test_indices, ]
test_pred <- pred[test_indices, ]


# Load Model
lm1 <- readRDS("TLS/intelimon models/March 2026 Intelimon Update/RMNP Intelimon Update/3. ROMO_East_RX/a. ROMO_East_RX_SubSev/ROMO_East_RX_SubSev.rda")


#look at the summary r2 values and RMSE & check model for colinearity
summary(lm1)
r2(lm1) 
rmse(lm1)
check_model(lm1) 

# Plot observed values vs. predicted
lm_out <- predict.lm(lm1, vars)
data<-data.frame(x=lm_out, y=pred$SubSev)

ggplot(data,aes(x,y)) +
  geom_point() +
  geom_smooth(method='lm', se=FALSE,) +
  theme_light() +
  labs(x='Predicted Values', y='Observed Values', title='Substrate Burn Severity') +
  theme(plot.title = element_text(hjust=0.5, size=20, face='bold') 
  ) 


# use the model to predict test data
test_out<- predict.lm(lm1,test_vars)
errors<- test_out - test_pred$SubSev
mae <- mean(abs(errors))
mse <- mean(errors^2)
rmse <- sqrt(mse)
rmse 

# Plot test observed values vs. test predicted
data2<- data.frame(x=test_out, y=test_pred$SubSev)
ggplot(data2,aes(x,y)) +
  geom_point() +
  geom_smooth(method='lm', se=FALSE,) +
  theme_light() +
  labs(x='Predicted Values', y='Observed Values', title='test') +
  theme(plot.title = element_text(hjust=0.5, size=20, face='bold') 
  ) 


```

## b. ROMO_East_RX_MeanScPct

model formula:
lm1 <- lm(formula = train_pred$MeanScPct ~    h_l3_tgi + LF_SLPD + s_lng_zero_per + TBA)

other top model variables: 
- LF_SLPD
- h_l2_std
- hr100_1000_l1_tgi

```{r}
library(tidyverse)
library(leaps)
library(performance)
library(ggplot2)

# Load pred / vars data
pred <- read.csv("TLS/intelimon models/March 2026 Intelimon Update/RMNP Intelimon Update/3. ROMO_East_RX/b. ROMO_East_RX_MeanScPct/ROMO_East_RX_MeanScPct_pred.csv")
vars <- read.csv("TLS/intelimon models/March 2026 Intelimon Update/RMNP Intelimon Update/3. ROMO_East_RX/b. ROMO_East_RX_MeanScPct/ROMO_East_RX_MeanScPct_vars.csv")

# Create training/test data
set.seed(123)
sample_size <- floor(0.25* nrow(vars))
test_indices <- sample(seq_len(nrow(vars)), size = sample_size)
train_vars <- vars[-test_indices, ]
test_vars <- vars[test_indices, ]
train_pred <- pred[-test_indices, ]
test_pred <- pred[test_indices, ]

# Load Model
lm1 <- readRDS("TLS/intelimon models/March 2026 Intelimon Update/RMNP Intelimon Update/3. ROMO_East_RX/b. ROMO_East_RX_MeanScPct/ROMO_East_RX_MeanScPct.rda")


#look at the summary r2 values and RMSE & check model for colinearity
summary(lm1)
r2(lm1) 
rmse(lm1)
check_model(lm1) 

# Plot observed values vs. predicted
lm_out <- predict.lm(lm1, vars)
data<-data.frame(x=lm_out, y=pred$MeanScPct)

ggplot(data,aes(x,y)) +
  geom_point() +
  geom_smooth(method='lm', se=FALSE,) +
  theme_light() +
  labs(x='Predicted Values', y='Observed Values', title='mean Scorch Percent') +
  theme(plot.title = element_text(hjust=0.5, size=20, face='bold') 
  ) 


# use the model to predict test data
test_out<- predict.lm(lm1,test_vars)
errors<- test_out - test_pred$MeanScPct
mae <- mean(abs(errors))
mse <- mean(errors^2)
rmse <- sqrt(mse)
rmse 

# Plot test observed values vs. test predicted
data2<- data.frame(x=test_out, y=test_pred$MeanScPct)
ggplot(data2,aes(x,y)) +
  geom_point() +
  geom_smooth(method='lm', se=FALSE,) +
  theme_light() +
  labs(x='Predicted Values', y='Observed Values', title='test') +
  theme(plot.title = element_text(hjust=0.5, size=20, face='bold') 
  ) 


```


## c. ROMO_East_RX_MeanScHt

model formula:
lm1 <- lm(formula = train_pred$MeanScHt ~ h_l1_tgi   +   h_GC_median +fuel0_3l1_median )

Other top model variables:
- USvol
- SDSHT
- h_l3_tgi
- ULAI
- LF_SLPD
- h_gc_tgi
- fine_l1_tgi

```{r}
library(tidyverse)
library(leaps)
library(performance)
library(ggplot2)

# Load pred / vars data
pred <- read.csv("TLS/intelimon models/March 2026 Intelimon Update/RMNP Intelimon Update/3. ROMO_East_RX/c. ROMO_East_RX_MeanScHt/ROMO_East_RX_MeanScHt_pred.csv")
vars <- read.csv("TLS/intelimon models/March 2026 Intelimon Update/RMNP Intelimon Update/3. ROMO_East_RX/c. ROMO_East_RX_MeanScHt/ROMO_East_RX_MeanScHt_vars.csv")

# Create training/test data
set.seed(123)
sample_size <- floor(0.2* nrow(vars))
test_indices <- sample(seq_len(nrow(vars)), size = sample_size)
train_vars <- vars[-test_indices, ]
test_vars <- vars[test_indices, ]
train_pred <- pred[-test_indices, ]
test_pred <- pred[test_indices, ]

# Load Model
lm1 <- readRDS("TLS/intelimon models/March 2026 Intelimon Update/RMNP Intelimon Update/3. ROMO_East_RX/c. ROMO_East_RX_MeanScHt/ROMO_East_RX_MeanScHt.rda")


#look at the summary r2 values and RMSE & check model for colinearity
summary(lm1)
r2(lm1) 
rmse(lm1)
check_model(lm1) 

# Plot observed values vs. predicted
lm_out <- predict.lm(lm1, vars)
data<-data.frame(x=lm_out, y=pred$MeanScHt)

ggplot(data,aes(x,y)) +
  geom_point() +
  geom_smooth(method='lm', se=FALSE,) +
  theme_light() +
  labs(x='Predicted Values', y='Observed Values', title='mean Scorch Height') +
  theme(plot.title = element_text(hjust=0.5, size=20, face='bold') 
  ) 


# use the model to predict test data
test_out<- predict.lm(lm1,test_vars)
errors<- test_out - test_pred$MeanScHt
mae <- mean(abs(errors))
mse <- mean(errors^2)
rmse <- sqrt(mse)
rmse 

# Plot test observed values vs. test predicted
data2<- data.frame(x=test_out, y=test_pred$MeanScHt)
ggplot(data2,aes(x,y)) +
  geom_point() +
  geom_smooth(method='lm', se=FALSE,) +
  theme_light() +
  labs(x='Predicted Values', y='Observed Values', title='test') +
  theme(plot.title = element_text(hjust=0.5, size=20, face='bold') 
  ) 


```






# 4. ROMO_West_Fuels

## a. ROMO_West_Fuels_TotalAll

model formula:
lm1 <- lm(formula = train_pred$TotalAll ~ h_MS_tgi +   h_OS_per  +  LF_FDist)


```{r}
library(tidyverse)
library(leaps)
library(performance)
library(ggplot2)

# Load pred / vars data
pred <- read.csv("TLS/intelimon models/March 2026 Intelimon Update/RMNP Intelimon Update/4. ROMO_West_Fuels/a. ROMO_West_Fuels_TotalAll/ROMO_West_Fuels_TotalAll_pred.csv")
vars <- read.csv("TLS/intelimon models/March 2026 Intelimon Update/RMNP Intelimon Update/4. ROMO_West_Fuels/a. ROMO_West_Fuels_TotalAll/ROMO_West_Fuels_TotalAll_vars.csv")

# Create training/test data
set.seed(123)
sample_size <- floor(0.2* nrow(vars))
test_indices <- sample(seq_len(nrow(vars)), size = sample_size)
train_vars <- vars[-test_indices, ]
test_vars <- vars[test_indices, ]
train_pred <- pred[-test_indices, ]
test_pred <- pred[test_indices, ]

# Load Model
lm1 <- readRDS("TLS/intelimon models/March 2026 Intelimon Update/RMNP Intelimon Update/4. ROMO_West_Fuels/a. ROMO_West_Fuels_TotalAll/ROMO_West_Fuels_TotalAll.rda")


#look at the summary r2 values and RMSE & check model for colinearity
summary(lm1)
r2(lm1) 
rmse(lm1)
check_model(lm1) 

# Plot observed values vs. predicted
lm_out <- predict.lm(lm1, vars)
data<-data.frame(x=lm_out, y=pred$TotalAll)

ggplot(data,aes(x,y)) +
  geom_point() +
  geom_smooth(method='lm', se=FALSE,) +
  theme_light() +
  labs(x='Predicted Values', y='Observed Values', title='Total Fuel Load') +
  theme(plot.title = element_text(hjust=0.5, size=20, face='bold') 
  ) 


# use the model to predict test data
test_out<- predict.lm(lm1,test_vars)
errors<- test_out - test_pred$TotalAll
mae <- mean(abs(errors))
mse <- mean(errors^2)
rmse <- sqrt(mse)
rmse 

# Plot test observed values vs. test predicted
data2<- data.frame(x=test_out, y=test_pred$TotalAll)
ggplot(data2,aes(x,y)) +
  geom_point() +
  geom_smooth(method='lm', se=FALSE,) +
  theme_light() +
  labs(x='Predicted Values', y='Observed Values', title='test') +
  theme(plot.title = element_text(hjust=0.5, size=20, face='bold') 
  ) 


```

