---
title: "DINO_Update"
author: "Christina Fossum"
date: "`r Sys.Date()`"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = "C:/Users/cfossum/OneDrive - DOI/Desktop/TLS repository")
```

# Models for UTDSP Intelimon Update

## Contents: 

1. BB_PC_Cheatgrass: Model for Cheatgrass percent coverin Bighorn Basin

2. BB_PC_NativeGrass: Model for native grass percent cover in Bighorn Basin

3. BB_PC_Shrub: Model for shrub percent cover





# 1. BB_PC_Cheatgrass

model formula:
lm1 <- lm(formula = train_pred$PC_BRTE ~ fuel0_3l1_kurt + shrubs_l1_kurt)

```{r}

library(tidyverse)
library(leaps)
library(performance)
library(ggplot2)

# Load pred / vars data
pred <- read.csv("TLS/intelimon models/DINO_Update/DINO_pred.csv") %>% select(-1)
vars <- read.csv("TLS/intelimon models/DINO_Update/DINO_vars.csv")%>% select(-1)


# Create training/test data
set.seed(123)
sample_size <- floor(0.3* nrow(vars))
test_indices <- sample(seq_len(nrow(vars)), size = sample_size)
train_vars <- vars[-test_indices, ]
test_vars <- vars[test_indices, ]
train_pred <- pred[-test_indices, ]
test_pred <- pred[test_indices, ]

# Load model
lm1<- readRDS("TLS/intelimon models/DINO_Update/BB_PC_Cheatgrass.rda")

#look at the summary r2 values and RMSE & check model for colinearity
summary(lm1)
r2(lm1) 
rmse(lm1)
check_model(lm1) 

# Plot observed values vs. predicted
lm_out <- predict.lm(lm1, vars)
data<-data.frame(x=lm_out, y=pred$PC_BRTE)

ggplot(data,aes(x,y)) +
  geom_point() +
  geom_smooth(method='lm', se=FALSE,) +
  theme_light() +
  labs(x='Predicted Values', y='Observed Values', title='Percent Cover: Cheatgrass') +
  theme(plot.title = element_text(hjust=0.5, size=20, face='bold') 
  ) 


# use the model to predict test data
test_out<- predict.lm(lm1,test_vars)
errors<- test_out - test_pred$PC_BRTE
mae <- mean(abs(errors))
mse <- mean(errors^2)
rmse <- sqrt(mse)
rmse 

# Plot test observed values vs. test predicted
data2<- data.frame(x=test_out, y=test_pred$PC_BRTE)
ggplot(data2,aes(x,y)) +
  geom_point() +
  geom_smooth(method='lm', se=FALSE,) +
  theme_light() +
  labs(x='Predicted Values', y='Observed Values', title='test') +
  theme(plot.title = element_text(hjust=0.5, size=20, face='bold') 
  ) 


```


# 2. BB_PC_NativeGrass

model formula:
lm1 <- lm(formula = train_pred$PC_NativeGrass ~ h_GC_kurt + fuel0_3l1_vari + hr0_10_l1_vari)

```{r}

library(tidyverse)
library(leaps)
library(performance)
library(ggplot2)

# Load pred / vars data
pred <- read.csv("TLS/intelimon models/DINO_Update/DINO_pred.csv") %>% select(-1)
vars <- read.csv("TLS/intelimon models/DINO_Update/DINO_vars.csv")%>% select(-1)


# Create training/test data
set.seed(123)
sample_size <- floor(0.2* nrow(vars))
test_indices <- sample(seq_len(nrow(vars)), size = sample_size)
train_vars <- vars[-test_indices, ]
test_vars <- vars[test_indices, ]
train_pred <- pred[-test_indices, ]
test_pred <- pred[test_indices, ]

# Load model
lm1<- readRDS("TLS/intelimon models/DINO_Update/BB_PC_NativeGrass.rda")

#look at the summary r2 values and RMSE & check model for colinearity
summary(lm1)
r2(lm1) 
rmse(lm1)
check_model(lm1) 

# Plot observed values vs. predicted
lm_out <- predict.lm(lm1, vars)
data<-data.frame(x=lm_out, y=pred$PC_NativeGrass)

ggplot(data,aes(x,y)) +
  geom_point() +
  geom_smooth(method='lm', se=FALSE,) +
  theme_light() +
  labs(x='Predicted Values', y='Observed Values', title='Percent Cover: Native Grass') +
  theme(plot.title = element_text(hjust=0.5, size=20, face='bold') 
  ) 


# use the model to predict test data
test_out<- predict.lm(lm1,test_vars)
errors<- test_out - test_pred$PC_NativeGrass
mae <- mean(abs(errors))
mse <- mean(errors^2)
rmse <- sqrt(mse)
rmse 

# Plot test observed values vs. test predicted
data2<- data.frame(x=test_out, y=test_pred$PC_NativeGrass)
ggplot(data2,aes(x,y)) +
  geom_point() +
  geom_smooth(method='lm', se=FALSE,) +
  theme_light() +
  labs(x='Predicted Values', y='Observed Values', title='test') +
  theme(plot.title = element_text(hjust=0.5, size=20, face='bold') 
  ) 


```


# 3. BB_PC_Shrub

model formula:
lm1 <- lm(formula = train_pred$PC_Shrub ~h_OS_per  +  vox_l1_tgi + fine_l1_vari )

```{r}

library(tidyverse)
library(leaps)
library(performance)
library(ggplot2)

# Load pred / vars data
pred <- read.csv("TLS/intelimon models/DINO_Update/DINO_pred.csv") %>% select(-1)
vars <- read.csv("TLS/intelimon models/DINO_Update/DINO_vars.csv")%>% select(-1)


# Create training/test data
set.seed(123)
sample_size <- floor(0.2* nrow(vars))
test_indices <- sample(seq_len(nrow(vars)), size = sample_size)
train_vars <- vars[-test_indices, ]
test_vars <- vars[test_indices, ]
train_pred <- pred[-test_indices, ]
test_pred <- pred[test_indices, ]

# Load model
lm1<- readRDS("TLS/intelimon models/DINO_Update/BB_PC_Shrub.rda")

#look at the summary r2 values and RMSE & check model for colinearity
summary(lm1)
r2(lm1) 
rmse(lm1)
check_model(lm1) 

# Plot observed values vs. predicted
lm_out <- predict.lm(lm1, vars)
data<-data.frame(x=lm_out, y=pred$PC_Shrub)

ggplot(data,aes(x,y)) +
  geom_point() +
  geom_smooth(method='lm', se=FALSE,) +
  theme_light() +
  labs(x='Predicted Values', y='Observed Values', title='Percent Cover: Native Grass') +
  theme(plot.title = element_text(hjust=0.5, size=20, face='bold') 
  ) 


# use the model to predict test data
test_out<- predict.lm(lm1,test_vars)
errors<- test_out - test_pred$PC_Shrub
mae <- mean(abs(errors))
mse <- mean(errors^2)
rmse <- sqrt(mse)
rmse 

# Plot test observed values vs. test predicted
data2<- data.frame(x=test_out, y=test_pred$PC_Shrub)
ggplot(data2,aes(x,y)) +
  geom_point() +
  geom_smooth(method='lm', se=FALSE,) +
  theme_light() +
  labs(x='Predicted Values', y='Observed Values', title='test') +
  theme(plot.title = element_text(hjust=0.5, size=20, face='bold') 
  ) 


```















